\begin{abstract}

Solving complex problems with large language models (LLMs) often requires selecting the best answer from multiple candidate responses (best-of-$N$), yet identifying the most reliable one remains challenging, particularly when correctness does not align with majority agreement.
Existing approaches based on voting or probability scoring rely on surface-level signals (e.g., frequency or token-level likelihood), often missing high-quality rare answers and ignoring the geometric structure of candidate representations, where correctness may emerge as semantic consistency rather than surface agreement.
To address these limitations, we introduce \textbf{Radial Consensus Score (RCS)}, a simple, efficient, and training-free method for best-of-$N$ selection. 
RCS models semantic consensus by computing a weighted Fréchet mean (semantic center)of answer embeddings and ranking candidates by their radial distance to this center. Importantly, RCS offers a flexible framework that supports multiple weighting schemes, enabling the integration of agreement signals and model confidence, where the position of center adapts accordingly.
% RCS further supports flexible weighting schemes to integrate agreement signals and model confidence, with the center adapting accordingly---for example, uniform averaging treats all candidates equally, while frequency or probability weighting shifts it toward more consistent or higher-confidence responses.
% RCS further provides a flexible framework that supports multiple weighting schemes, enabling the integration of agreement signals and model confidence. 
% The position of this center adapts to different consensus signals through the choice of weighting scheme: uniform averaging treats all candidates equally, while frequency or probability weighting shifts the center toward more consistent or higher-confidence responses.
% RCS provides a general framework that supports multiple weighting schemes, enabling flexible integration of agreement signals and model confidence while remaining fully applicable in black-box settings.
% In practice, the consensus center can be instantiated via standard embedding space averaging or other weighted aggregation methods, depending on the available signals.
Extensive experiments across various benchmarks covering short-form QA and long-form reasoning tasks, and five open-weight models, demonstrate that our metrics consistently outperform strong baselines with advantages growing as $N$ increases.
Moreover, RCS also serves as an effective drop-in replacement for majority voting in multi-agent debate and exhibits strong robustness in black-box scenarios.
These results highlight geometric consensus as an applicable and scalable principle for reliable answer selection beyond majority voting.
\end{abstract}